Hysteresis in human computation:\\ how one task affects another

### 2014-12-04 docs/publications/2014.09_SCIENCE/man.tex
catch more typos and update plot script

In all treatments, workers were shown images one at a time, and provided five 
labels for each.  From the perspective of the worker, there was no
distinction or interruption between the initial and test tasks.  We used
workers' unique identifiers to ensure that they only participated in a 
one treatment of a single experiment.

@@

To couch intertask effects in a practical yet general setting, we imagine
that, after a worker performs a task, her response is used to make some 
kind of yes-or-no decision, $\mathcal{D}$.  For example, this decision 
could be whether or not to flag a certain comment
below a news article as inappropriate; but for now, we leave 
$\mathcal{D}$ unspecified.  Since workers' responses may
vary, repeating the same task might lead to different outcomes for 
$\mathcal{D}$.  This decision not only models the reality that responses are
actually used---introducing $\mathcal{D}$ allows us to define of \textit{bias}
in a practical yet general way. We shall define bias in terms of the
change in the probability of a positive outcome from $\mathcal{D}$, induced 
by prior exposures.

@@

Next, we looked at the change in the frequency of occurrence of particular 
words.  Surprisingly, the word ``food'' was the \textit{most suppressed} word
for workers in \textit{*:food} across all treatment pairs (Table S1 shows the
top five most promoted and suppressed words for each treatment pair).
Initially, this seems at odds with the fact that (at least in \textit{exp1})
the \textit{*:food} treatments had more food references than their conjugate
treatments.  But taken together, these facts indicate that workers are opting 
to use more specialized food-references in place of ``food''.  
Indeed, when we look at the total number of unique
words (lexical richness) used to refer to food, we found that 
a significantly richer lexicon was used by all \textit{*:food} treatments,
except in \textit{frame2} (see Fig.~\ref{fig:specificity}B).  This effect 
was especially pronounced in \textit{task1} and 
\textit{task2}.

@@

Workers exposed to images containing food will be (positively) primed, 
as memories, concepts, and vocabulary related to food are activated.  
However, we argue that as the worker labels successive images containing food,
the basic 
fact that an image contains food will not appear to be salient, since it 
does not serve to distinguish one image from another.  Thus, the most 
generic references to that fact, such as the label ``food'', 
are suppressed, as by negative priming, while more specialized
references are activated.  Whether workers in \textit{*:food} treatments
produce an excess of food references overall depends in part on the balance 
of these factors.  The reduction in food references from workers in 
\textit{task2:food} relative
to those in \textit{task2:cult} may also be related to the novelty that 
workers in 
\textit{task2:cult} experience when they first encounter food in the test 
tasks.

@@

\paragraph{Recommendations and future work.} 
Based on our observations throughout, we submit the following 
recommendations for the consideration of people designing 
any microtask, citizen science, or crowdsourcing initiative:
\begin{enumerate}
	\item{
		Require active feedback to emphasize important details about the 
		tasks, as we did in \textit{echo1}.  Similarly, use training tasks 
		rather than only examples.
	}
	\item{
		Although ``clean'' examples may be useful to provide clarity,
		use training tasks that contain the same kinds of noise or otherwise
		distracting features as are expected in the real tasks.
		This exposure will negatively prime the worker against non-salient 
		features.
	}
	\item{
		\textit{Validation tasks}, for which the correct output is known,
		are often used as a quality control measure.
		Treat these as \textit{calibration tasks}, and use them to regulate
		the focus of the worker.
		Feedback, for both correct and incorrect responses may further 
		reinforce the desired competency. \label{item:valid}
	}
	\item{
		A special case of item~\ref{item:valid} applies when workers are 
		required to detect rare occurrences.
		For example, in the detection of pre-ictal (pre-seizure) EEG traces, 
		most traces are negative (\#).  In the absence of positive 
		examples, the target concept may degrade or drift.
		Draw calibration tasks from the underrepresented class,
		and provide feedback, to help sustain the worker's target concept.
	}
	\item{
		For more precise classifications, it may help to begin with
		an initial coarse classification.  For example, in detecting 
		inappropriate comments, provide the choices   
		\texttt{okay}, \texttt{inappropriate}, and \texttt{borderline},
		and incentivise inter-worker agreement
		Next, batch the \texttt{borderline}.  Coding this more homogeneous
		borderline set will help workers to calibrate to the finer 
		distinctions necessary for the task.
	}
\end{enumerate}

@@

\paragraph{Conclusion.}
We have shown that intertask effects are stronger than framing effects, and
as strong as the effects induced when framing is combined with reiteration.
Intertask and framing effects act on workers' vocabulary in subtle ways.  
Both increase the richness of workers' vocabulary in reference a target
concept, with a more pronounced effect induced by task exposure.
Both exposures also increase the specialization of workers' vocabulary in 
reference to the target concept.  Framing, and other aspects of task design 
have received substantial methodological scrutiny; our results show that 
intertask effects, which were overlooked until now, merit similar 
attention.  Although we found that intertask effects can be a major 
source of bias, they might be used to tune workers' attention and 
specificity relative to
a target concept.  Microtask work, citizen science, and other crowdsourcing
initiatives, as well task requiring expert opinions, often rely on repetitive 
human judgments.  Therefore, such applications all stand to benefit from the 
findings and suggestions presented here.

>>>>

In all treatments, workers were shown images one at a time, and provided five 
labels for each.  From the perspective of the worker, there was no
distinction or interruption between the initial and test tasks.  We used
workers' unique identifiers to ensure that they only participated in
one treatment of a single experiment.

@@

To couch intertask effects in a practical yet general setting, we imagine
that, after a worker performs a task, her response is used to make some 
kind of yes-or-no decision, $\mathcal{D}$.  For example, this decision 
could be whether or not to flag a certain comment
below a news article as inappropriate; but for now, we leave 
$\mathcal{D}$ unspecified.  Since workers' responses may
vary, repeating the same task might lead to different outcomes for 
$\mathcal{D}$.  This decision not only models the reality that responses are
actually used---introducing $\mathcal{D}$ allows us to define \textit{bias}
in a practical yet general way. We shall define bias in terms of the
change in the probability of a positive outcome from $\mathcal{D}$, induced 
by prior exposures.

@@

Next, we looked at the change in the frequency of occurrence of particular 
words.  Surprisingly, the word ``food'' was the \textit{most suppressed} word
for workers in \textit{*:food} across all treatment pairs (Table S1 shows the
top five most promoted and suppressed words for each treatment pair).
Initially, this seems at odds with the fact that (at least in \textit{exp1})
the \textit{*:food} treatments had more food references than their conjugate
treatments.  But taken together, these facts indicate that \textit{*:food}
workers are opting 
to use more specialized food-references in place of ``food''.  
Indeed, when we look at the total number of unique
words (lexical richness) used to refer to food, we found that 
a significantly richer lexicon was used by all \textit{*:food} treatments,
except in \textit{frame2} (see Fig.~\ref{fig:specificity}B).  This effect 
was especially pronounced in \textit{task1} and 
\textit{task2}.

@@

Workers exposed to images containing food will be (positively) primed, 
as memories, concepts, and vocabulary related to food are activated.  
However, we argue that as the worker labels successive images containing food,
the basic 
fact that an image contains food will not appear to be salient, since it 
does not serve to distinguish one image from another.  Thus, the most 
generic references to that fact, such as the label ``food'', 
are suppressed, as by negative priming, while more specialized
references are activated.  Whether workers in \textit{*:food} treatments
produce an excess of food references overall depends in part on the balance 
of these factors.  The reduction in food references from workers in 
\textit{task2:food} relative to \textit{task2:cult}, seen in in 
Fig.~\ref{fig:specificity}A, may also be
related to the novelty that workers in 
\textit{task2:cult} experience when they first encounter food in the test 
tasks.

@@

\paragraph{Recommendations and future work.} 
Based on our observations throughout, we submit the following 
recommendations for the consideration of people designing 
any microtask, citizen science, or crowdsourcing initiative:
\begin{enumerate}
	\item{
		Require active feedback to emphasize important details about the 
		tasks, as we did in \textit{echo1}.  Similarly, use training tasks 
		rather than only examples.
	}
	\item{
		Although ``clean'' examples may be useful to provide clarity,
		use training tasks that contain the same kinds of noise or otherwise
		distracting features as are expected in the real tasks.
		This exposure will negatively prime the worker against non-salient 
		features.
	}
	\item{
		\textit{Validation tasks}, for which the correct output is known,
		are often used as a quality control measure.
		Treat these as \textit{calibration tasks}, and use them to regulate
		the focus of the worker.
		Feedback, for both correct and incorrect responses may further 
		reinforce the desired competency. \label{item:valid}
	}
	\item{
		A special case of item~\ref{item:valid} applies when workers are 
		required to detect rare occurrences.
		For example, in the detection of pre-ictal (pre-seizure) EEG traces, 
		most traces are negative (\#).  In the absence of positive 
		examples, the target concept may degrade or drift.
		Draw calibration tasks from the underrepresented class,
		and provide feedback, to help sustain the worker's target concept.
	}
	\item{
		For more precise classifications, it may help to begin with
		an initial coarse classification.  For example, in detecting 
		inappropriate comments, provide the choices   
		\texttt{okay}, \texttt{inappropriate}, and \texttt{borderline},
		and incentivise inter-worker agreement.  Next, have workers 
		code the \texttt{borderline} cases, but without providing the
		\texttt{borderline} code.  Coding this more homogeneous
		set will help workers to calibrate to the finer features that are 
		salient in distinguishing borderline cases.
	}
\end{enumerate}

@@

\paragraph{Conclusion.}
We have shown that intertask effects are stronger than framing effects, and
as strong as the effects induced when framing is combined with reiteration.
Intertask and framing effects act on workers' vocabulary in subtle ways.  
Both increase the richness of workers' vocabulary in reference a target
concept, with a more pronounced effect induced by task exposure.
Both exposures also increase the specialization of workers' vocabulary in 
reference to the target concept.  Framing, and other aspects of task design 
have received substantial methodological scrutiny; our results show that 
intertask effects, which were overlooked until now, merit similar 
attention.  Although we found that intertask effects can be a major 
source of bias, they might be used to tune workers' attention and 
specificity relative to
a target concept.  Microtask work, citizen science, and other crowdsourcing
initiatives, as well tasks requiring expert opinions, often rely on repetitive
human judgments.  Such applications all stand to benefit from the 
findings and suggestions presented here.
